GMSE: an R package for generalised management strategy evaluation

### 2018-04-13 notebook/ms/SI1.Rmd
SI changes

Options for adjusting

| GMSE argument       | Default | Description |
|---------------------|---------|-------------|
| `ga_popsize`        | 100     | The number of individuals in the population temporary simulated during a single run of the genetic algorithm. |
| `ga_mingen`         | 40      | The minimum number of iterations that a genetic algorithm will run before settling on an agent's strategy. |
| `ga_seedrep`        | 20      | The number of individuas in the population to be initiaised with the current agent's strategy (e.g., from a previous time step in the broader GMSE simulation), as opposed to being initialised with random strategies. | 
| `ga_sampleK`        | 20      | For the tournament step of the genetic agorithm, how many strategies are selected at random from the larger population (with replacement) to be included a the tournament. |
| `ga_chooseK`        | 2       | Four the tournament step of the genetic agorithm, how many strategies are selected as winners of the tournament, to be included in the next iteration. |
| `ga_mutation`       | 0.1     | The mutation rate of any action in an agent's strategy |
| `ga_crossover`      | 0.1     | The crossover rate of any action in an agent's strategy; crossover events occur with a different randomly selected strategy in the population. |
| `ga_converge_crit`  | 1       | The percent increase in strategy fitness from one iteration to the next below which the convergence criteria is satisfied. Iterations wil continue as long as fitness increase is above this convergence criteria. |
| `group_think`       | FALSE   | Whether or not all users (i.e., not including the manager) have identical strategies. If TRUE, then one genetic algorithm will be run and applied to a users >

>>>>

Options for adjusting genetic algorithm parameter values in `gmse` and `gmse_apply` are shown below.

| GMSE argument       | Default | Description                                  |
|---------------------|---------|----------------------------------------------|
| `ga_popsize`        | 100     | The number of individuals in the population temporary simulated during a single run of the genetic algorithm. |
| `ga_mingen`         | 40      | The minimum number of iterations that a genetic algorithm will run before settling on an agent's strategy. |
| `ga_seedrep`        | 0       | The number of individuas in the population to be initiaised with the current agent's strategy (e.g., from a previous time step in the broader GMSE simulation), as opposed to being initialised with random strategies. | 
| `ga_sampleK`        | 20      | For the tournament step of the genetic agorithm, how many strategies are selected at random from the larger population (with replacement) to be included a the tournament. |
| `ga_chooseK`        | 2       | Four the tournament step of the genetic agorithm, how many strategies are selected as winners of the tournament, to be included in the next iteration. |
| `ga_mutation`       | 0.1     | The mutation rate of any action in an agent's strategy |
| `ga_crossover`      | 0.1     | The crossover rate of any action in an agent's strategy; crossover events occur with a different randomly selected strategy in the population. |
| `ga_converge_crit`  | 1       | The percent increase in strategy fitness from one iteration to the next below which the convergence criteria is satisfied. Iterations wil continue as long as fitness increase is above this convergence criteria. |
| `group_think`       | FALSE   | Whether or not all users (i.e., not including the manager) have identical strategies. If TRUE, then one genetic algorithm will be run and applied to all users. |

Given the heuristic goals of the genetic algorithm to mimic the goal-oriented behaviour of agents, default parameters are typically sufficient for agent decision making. Key parameters can be adjusted if more precision in decision making is desired, but these adjustments will come at a cost of simulation efficiency. For example, increasing `ga_popsize` or `ga_mingen`, or decreasing `ga_converge_crit`, might fine tune strategies more effectively, but this will cause the genetic algorithm to take longer every time that it is run, ultimately slowing down GMSE simulations. Alternativey, setting `group_think = TRUE` will greatly speed up GMSE simulations when many users are being simulated, but this comes at the cost of among-user variation in decision making. Overall, we recommend first using default values in the genetic algorithm before exploring how other parameter value options affect simulation dynamics; for a more general discussion about selecting parameter values in genetic algorithms, see @Hamblin2013.

### 2018-04-13 notebook/ms/SI3.Rmd
SI changes

To demonstrate how GMSE can be used to understand conflict in social-ecological systems, we build upon the [example of resource management](https://www.biorxiv.org/content/biorxiv/early/2017/11/17/221432.full.pdf#page=5) in the main text. We consider a protected population of waterfowl that exploits agricultural land and is therefore a source of conservation conflict [e.g., @Fox2017a; @Mason2017; @Tulloch2017]. As in the main text example, the objective of managers is to keep waterfowl at a target abundance, while the objective of farmers is to maximise agricultural production on their landscape. Here we consider a more complex simulation to reflect a scenario likely to occur in a real social-ecological system. In contrast to the main text, we consider a population of waterfowl that is abundant enough to substantially decrease yield. We consider a community of 100 farmers, each of which has a farm of 50 landscape cells, which could be interpreted as hectares of arable cropland. We additionally include landscape cells that are not owned by farmers using the `public_land` argument; these cells can be interpreted as public land on which waterfowl might forage, but which do not affect farmers' agricultural production. We further assume an intrinsic growth rate of 0.5, and other parameters as provided by literature for geese (fix this).

```{r, echo = TRUE}
sim <- gmse(land_ownership = TRUE, stakeholders = 100, observe_type = 0, 
            res_death_K = 20000, manage_target = 10000, RESOURCE_ini = 8000, 
            user_budget = 1000, manager_budget = 8000, res_consume = 1, 
            scaring = FALSE, plotting = FALSE, public_land = 0.5);
```

@@

```{r, eval = FALSE}
plot_gmse_results(res = sim$resource, obs = sim$observation, land = sim$land, 
                  agents = sim$agents, paras = sim$paras, ACTION = sim$action, 
                  COST = sim$cost);
```

>>>>

To demonstrate how GMSE can be used to understand conflict in social-ecological systems, we build upon the [example of resource management](https://www.biorxiv.org/content/biorxiv/early/2017/11/17/221432.full.pdf#page=5) in the main text. We consider a protected population of waterfowl that exploits agricultural land and is therefore a source of conservation conflict [e.g., @Fox2017a; @Mason2017; @Tulloch2017]. As in the main text example, the objective of managers is to keep waterfowl at a target abundance, while the objective of farmers is to maximise agricultural production on their landscape. Here we consider a more complex simulation to reflect a scenario likely to occur in a real social-ecological system. In contrast to the main text, we consider a population of waterfowl that is abundant enough to substantially decrease yield. We consider a community of 20 farmers, each of which has a farm of 50 landscape cells, which could be interpreted as hectares of arable cropland. We additionally include landscape cells that are not owned by farmers using the `public_land` argument; these comprise 40 percent of cells and can be interpreted as public land on which waterfowl might forage, but which do not affect farmers' agricultural production. We further assume an intrinsic growth rate of 0.5, and other parameters as provided by literature for geese (fix this).

For simplicity, here we assess confict using only the `gmse` function to show how parameter values can be set to provide useful results. All of the analysis shown would also be possibe using the more detailed, but also more flexible, `gmse_apply` function.

```{r, echo = TRUE}
sim <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
            manage_target = 70000, RESOURCE_ini = 63000, plotting = FALSE, 
            stakeholders = 20, land_ownership = TRUE, land_dim_1 = 50, 
            land_dim_2 = 50, public_land = 0.4, scaring = FALSE, lambda = 0.275, 
            remove_pr = 0.122, time_max = 20, res_death_type = 3, ga_seedrep = 0, 
            ga_mingen = 100, res_consume = 0.02);
```

@@

```{r, eval = TRUE, fig.height = 8}
plot_gmse_results(res = sim$resource, obs = sim$observation, land = sim$land, 
                  agents = sim$agents, paras = sim$paras, ACTION = sim$action, 
                  COST = sim$cost);
```

Look at the conflict

```{r, eval = TRUE, fig.height = 7, fig.width = 6}
plot_gmse_effort(sim$agents, sim$paras, ACTION = sim$action, COST = sim$cost);
```

Now try scaring to see what happens.

```{r, echo = TRUE, fig.height = 8}
sim <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
            manage_target = 70000, RESOURCE_ini = 63000, plotting = FALSE, 
            stakeholders = 20, land_ownership = TRUE, land_dim_1 = 50, 
            land_dim_2 = 50, public_land = 0.4, scaring = TRUE, lambda = 0.275, 
            remove_pr = 0.122, time_max = 20, res_death_type = 3, ga_seedrep = 0, 
            ga_mingen = 100, res_consume = 0.02);
```

```{r, eval = TRUE, fig.height = 8}
plot_gmse_results(res = sim$resource, obs = sim$observation, land = sim$land, 
                  agents = sim$agents, paras = sim$paras, ACTION = sim$action, 
                  COST = sim$cost);
```

Look at the conflict

```{r, eval = TRUE, fig.height = 7, fig.width = 6}
plot_gmse_effort(sim$agents, sim$paras, ACTION = sim$action, COST = sim$cost);
```
